---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# gRs

<!-- badges: start -->
<!-- badges: end -->

The goal of gRs is to process and analyse surfacewater and groundwater data from esdat.
Tha initial goal is to implement Mann-Kendall analysis on an entire dataset at once and produce data visualisation of those trends.

## Installation

You can install the development version of gRs from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("jenright-git/gRs")
```

## Example

This is a basic example which shows you how to solve a common problem:

Import data from an excel file exported directly from esdat.

```{r example}
library(gRs)

#import data
#gRs_data <- data_processor(("gRs_data.xlsx"))

```

Perform Mann-Kendall trend test
```{r}
gRs::gRs_data %>% 
  mann_kendall_test() %>% 
  head(10) %>% 
  gt::gt()
```

```{r}
gRs::gRs_data %>% 
  mann_kendall_test() %>% 
  mann_kendall_heatmap()
```

