---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  warning = FALSE,
  message=FALSE
)
```

# gRs

<!-- badges: start -->
<!-- badges: end -->

The goal of gRs is to process and analyse surface water, groundwater and soil/sediment data from esdat and EQuIS.
Tha initial goal is to implement Mann-Kendall analysis on an entire dataset at once and produce data visualisation of those trends.

`data_processor()` reads chemistry exports (EQuIS Analytical Results II, esdat `LChem1_Chemistry` for liquids and `SChem1_Chemistry` for soils, and esdat `Chemistry List`) and gauging reports (EQuIS `Water Levels II` and esdat gauging reports), normalising each to a shared set of column names. The format is detected from the worksheet contents, so no extra arguments are needed.

Soil exports report each analyte twice - once as a solid-phase concentration (mg/kg) and once as a leachate concentration (mg/L or ug/L). Only the solid-phase results are returned by default; pass `result_type = "LEACHED_REG"` for the leachate results, or `result_type = "all"` for both.

## Installation

You can install the development version of gRs from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("jenright-git/gRs")
```

## Example

This is a basic example which shows you how to solve a common problem:

Import data from an excel file exported directly from esdat.

```{r example, warning=FALSE, message=FALSE}
library(gRs)
library(tidyverse)

establish_plotting_variables(gRs::gRs_data)

#import data
#gRs_data <- data_processor(("gRs_data.xlsx"))

```

Perform Mann-Kendall trend test
```{r, warning=FALSE, message=FALSE}
gRs::gRs_data %>% 
  gRs::mann_kendall_test() %>% 
  select(-data) %>% 
  head(5)
```

Visualise trends with a heatmap
```{r, fig.width=15, fig.height=12, warning=FALSE, message=FALSE}
gRs::gRs_data %>% 
  mann_kendall_test() %>% 
  mann_kendall_heatmap(width=18)
```

Plot the increasing trends.
```{r}
gRs::gRs_data %>% 
  mann_kendall_test() %>% 
  filter(trend == "Increasing") %>% 
  unnest(data) %>% 
  timeseries_plot(date_break = "2 month", date_label = "%b")+
  facet_wrap(~chem_name, scales="free_y")
```

