---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  warning = FALSE,
  message = FALSE
)
```

# gRs

<!-- badges: start -->
<!-- badges: end -->

An R package for reading, analysing and visualising esdat and EQuIS environmental data.

## Installation

``` r
# install.packages("devtools")
devtools::install_github("jenright-git/gRs")

```

## Reading data

`data_processor()` reads excel exports from the esdat data view or grid reports from EQuIS. These include soil and water chemistry as well as water level exports.

``` r
chem <- data_processor("davLChem1_Chemistry.xlsx")
```

`action_level_processor()` reads an esdat action level export.

``` r
al <- action_level_processor("ANZG 95 Marine.xlsx", name = "ANZG 95% Marine")
```

`gRs_data` is bundled example data, in the shape `data_processor()` returns.

```{r example-data}
library(gRs)
library(dplyr)
library(gt)

gRs_data |>
  select(date, monitoring_round, location_code, chem_name, prefix,
         concentration, output_unit, detect_flag) |>
  head(6) 
```

## Preparing results

`select_max_concentration()` collapses field and interlab duplicates to the highest result.

``` r
chem <- select_max_concentration(chem)
```

`half_lor()` substitutes a fraction of the LOR for every non-detect.

```{r half-lor}
gRs_data |>
  filter(chem_name == "Chlorophyll a", !is.na(prefix)) |>
  half_lor(lor_multiplier = 0.5) |>
  select(date, location_code, prefix, concentration, detect_flag) |>
  head(4) 
```

```{r hand-built-guidelines, include = FALSE}
# A small guideline set standing in for an esdat export, so the examples below
# have something to join against.
al <- tibble::tribble(
  ~chem_name,               ~chem_code,           ~criteria, ~criteria_unit,
  "Total Nitrogen (as N)",  "7727-37-9",              0.3,   "mg/L",
  "Phosphorus",             "7723-14-0",              0.03,  "mg/L",
  "Ammonia (as N)",         "7664-41-7 (as N)",       0.015, "mg/L",
  "Nitrite + Nitrate as N", "NO3+NO2 (as N)",         0.015, "mg/L",
  "Chlorophyll a",          "42617-16-3",             4,     "ug/L",
  "E. Coli",                "E.Coli",               150,     "CFU/100mL",
  "Dissolved Reactive Phosphorus as P (Orthophosphate as P)",
                            "7723-14-0_Reac_AsP",     0.005, "mg/L"
) |>
  mutate(
    criteria_name  = "ANZG 95% Estuarine",
    filtered       = chem_code == "7723-14-0_Reac_AsP",
    total          = !filtered,
    leached        = FALSE,
    matrix_code    = NA_character_,
    criteria_basis = NA_character_
  )
```

`join_action_levels()` puts a guideline against every result and flags the exceedances.

```{r join-guidelines}
compared <- join_action_levels(gRs_data, al, match_matrix = FALSE)

compared |>
  filter(chem_name == "Chlorophyll a") |>
  select(location_code, date, concentration, output_unit,
         criteria, exceedance_ratio, exceedance) |>
  head(4) 
```

## Summary tables

`analyte_summary()` summarises a monitoring round, one row per analyte, and `create_gt()` formats it for a report.

```{r analyte-summary}
analyte_summary(compared, quiet = TRUE) 
```

`exceedance_summary()` returns the results behind those exceedance counts.

```{r exceedance-summary}
exceedance_summary(compared, quiet = TRUE) |>
  select(location_code, chem_name, concentration, output_unit,
         criteria, exceedance_ratio) |>
  head(8)
```

`min_max_locations()` returns the highest and lowest result for each analyte, with its location.

```{r min-max-locations}
min_max_locations(gRs_data, n_max = 1, n_min = 1, quiet = TRUE) |>
  select(chem_name, extreme, location_code, prefix,
         concentration, output_unit) |>
  head(8) 
```

`summary_stats()` summarises the whole record, one row per location and analyte.

```{r summary-stats}
summary_stats(compared, include_criteria = TRUE) |>
  filter(chem_name == "Ammonia (as N)") |>
  select(location_code, n_samples, n_detects, pct_detects,
         min, mean, max, p95, criteria, exceedance_count) 
```


## Historical range

`historical_range()` compares the latest round against the record behind it, one row per location and analyte.

```{r historical-range}
compared |>
  filter(chem_name == "Phosphorus") |>
  historical_range(quiet = TRUE) 
```

`keep = "new_max"` cuts it to the results that set a new maximum.

```{r historical-range-new-max}
historical_range(compared, keep = "new_max", quiet = TRUE) |>
  select(chem_name, location_code, hist_max_prefix, hist_max_conc,
         current_prefix, current_conc) 
```

## Trend analysis

`mann_kendall_test()` runs a Mann-Kendall test on every location and analyte at once.

```{r mann-kendall}
trends <- mann_kendall_test(gRs_data)

trends |>
  select(location_code, chem_name, p_value, tau_statistic, trend) |>
  head(5) 
```

`mann_kendall_heatmap()` plots the whole grid.

```{r trend-heatmap, fig.width = 15, fig.height = 12}
mann_kendall_heatmap(trends, width = 18)
```

`mka_to_excel()` writes the same grid to a formatted workbook - analytes across
the top, trends in the cells, coloured by direction - with a second sheet
explaining the categories.

```{r mka-excel, eval = FALSE}
mka_to_excel(trends, save_path = "output/MKA Trend Summary.xlsx")

# Recolour one category and leave the rest alone
mka_to_excel(trends, trend_colours = c(Increasing = "#E06666"))

# Group the wells by monitoring zone, read out of the nested data
mka_to_excel(trends, include_zone = TRUE)
```

## Plotting

`timeseries_plot()` facets by analyte and colours by location.

```{r increasing-trends}
library(tidyr)

trends |>
  filter(trend == "Increasing") |>
  unnest(data) |>
  timeseries_plot(date_break = "2 month", date_label = "%b")
```

One analyte at two locations, with `criteria_col` drawing the guideline.

```{r two-locations, fig.width = 8, fig.height = 5, fig.alt = "Monthly ammonia as N results at LOC_01 and LOC_04, each drawn as a coloured line, against a dashed horizontal line marking the 0.015 mg/L guideline."}
compared |>
  timeseries_plot(
    filter_analyte = "Ammonia (as N)",
    filter_location = c("LOC_01", "LOC_04"),
    criteria_col = criteria,
    date_break = "1 month",
    date_label = "%b-%y",
    plot_title = "Ammonia (as N)"
  )
```

Full documentation for every function is at <https://jenright-git.github.io/gRs/>.
